### 2018-10-17 Docs/Marami2018/Communication/marami2018.tex
slides

\documentclass[english,11pt]{beamer}

\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Proba}{\mathbb{P}}

\newcommand{\Covb}[2]{\ensuremath{\Cov\!\left[#1,#2\right]}}
\newcommand{\Eb}[1]{\ensuremath{\E\!\left[#1\right]}}
\newcommand{\Pb}[1]{\ensuremath{\Proba\!\left[#1\right]}}
\newcommand{\Varb}[1]{\ensuremath{\Var\!\left[#1\right]}}

% norm
\newcommand{\norm}[1]{\| #1 \|}

\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}

\usepackage{mathptmx,amsmath,amssymb,graphicx,bibentry,bbm,babel,ragged2e}

\makeatletter

\newcommand{\noun}[1]{\textsc{#1}}
\newcommand{\jitem}[1]{\item \begin{justify} #1 \end{justify} \vfill{}}
\newcommand{\sframe}[2]{\frame{\frametitle{#1} #2}}

\newenvironment{centercolumns}{\begin{columns}[c]}{\end{columns}}
%\newenvironment{jitem}{\begin{justify}\begin{itemize}}{\end{itemize}\end{justify}}

\usetheme{Warsaw}
\setbeamertemplate{footline}[text line]{}
\setbeamercolor{structure}{fg=purple!50!blue, bg=purple!50!blue}

\setbeamersize{text margin left=15pt,text margin right=15pt}

\setbeamercovered{transparent}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{multirow}

\makeatother

\begin{document}

\title{A multi-dimensional percolation approach to characterize sustainable mega-city regions}

\author{J.~Raimbault$^{1,2,\ast}$\\
\texttt{juste.raimbault@polytechnique.edu}
}

\institute{$^{1}$Complex Systems Institute, Paris, UPS CNRS 3611 ISC-PIF\\
$^{2}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
}

\date{MARAMI 2018\\\smallskip
Avignon\\\smallskip
October 18th 2018
}

\frame{\maketitle}

\sframe{Morphologies of networks and territories}{

% Keywords : }\textit{Road network; multilayer percolation; mega-city region}

% - intro : LGV ? Avignon : travaux Claire ! - truc avec les Papes ? 
% https://www.persee.fr/doc/anami_0003-4398_1985_num_97_172_2228_t1_0446_0000_3
% https://www.persee.fr/doc/mefr_0223-5110_1984_num_96_1_2762

% -> visu OSM : see Metropolitan region ; plus positioning of Avignon ; Q of sustainibility ?

\centering

	\includegraphics[width=1.2\textwidth]{figures/roadssouth.png}

	\footnotesize
\textit{Source: OpenStreetMap}

}

\sframe{Characterizing Road networks}{

%The structure of road networks both translates its past growth dynamics and has a significant impact on the sustainability of territories it irrigates.

% here extracts from Claire's work
% Lagesse, C., Bordin, P., & Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. -> nice picture Avignon

\textit{Multiple dimensions to characterize road networks}

\medskip

{
\centering
\includegraphics[width=\textwidth,height=0.7\textheight]{figures/avignon_claire.png}
}

\footnotesize

Lagesse, C., Bordin, P., \& Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. \cite{lagesse2015spatial}

}

\sframe{Network percolation}{

% A method to characterize topologies of these spatial networks is network percolation. Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling} and to the analysis of street networks for example to extract endogenous urban regions \citep{arcaute2016cities} or to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.

% from monodimensional to multidimensional

% similar to 
%Cottineau, C., Finance, O., Hatna, E., Arcaute, E., & Batty, M. (2018). Defining urban clusters to detect agglomeration economies. Environment and Planning B: Urban Analytics and City Science, 2399808318755146.

%

\justify

\textbf{Network percolation: } \textit{progressive occupation/connection of nodes of a network} \cite{callaway2000network}

\bigskip

Application to the study of cities:

\begin{itemize}
	\item modeling urban growth \cite{makse1998modeling}
	\item endeogenous determination of regions \cite{arcaute2016cities}
	\item characterization of spatial point patterns \cite{huynh2018characterisation}
\end{itemize}

\bigskip

\textit{Towards complementary dimensions to condition road network percolation}

$\rightarrow$ similar to \cite{cottineau2018defining} to define urban areas

}

\sframe{Multidimensional percolation}{

% Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.

% - why the relation between form and function, aka pop and network in our view is crucial
% - take into account in percolation
% - potential application : endogenous sustainable urban entities

\justify

$\rightarrow$ Need to combine morphological and functional dimensions of cities \cite{burger2012form}

\medskip

$\rightarrow$ Interaction between networks and territories to capture the form-function link \cite{raimbault2018caracterisation}; potential application to sustainability of urban systems

% research question

\bigskip
\bigskip

\textbf{Research objective : } \textit{Investigate a multi-dimensional percolation of territorial networks taking into account urban morphology and road network topology; endogenous characterization of urban regions.}

}

% \section{Method}

\sframe{Multilayer percolation}{

%This communication addresses such a gap by introducing a multi-dimensional percolation heuristic, which is analog to multilayer network percolation \citep{boccaletti2014structure}. Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold.

\textit{Multi-dimensional network percolation heuristic, similar to multilayer percolation} \cite{boccaletti2014structure}

\bigskip

\centering
\includegraphics[width=0.8\textwidth]{figures/principle.pdf}

\bigskip

\raggedright

\textbf{Parameters: } percolation radius $r_0$, percolation thresholds $\theta_i$ for each layer

}

\sframe{Empirical data and variables}{

%We apply the heuristic to urban morphology and road network topology measures in Europe. More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}.

\justify

\textit{Territorial indicators computed for Europe by \cite{raimbault2018urban}}

% here add maps to better visualize the spatial fields

\medskip

\includegraphics[width=0.49\textwidth]{figures/indics_morpho_discrquantiles.png}
\includegraphics[width=0.49\textwidth]{figures/indics_network_en_areasize100_offset50_factor0_5.png}

\medskip

\textbf{Population distribution morphology} and \textbf{Network topology} (betweenness, closeness, clustering, efficiency, \ldots) computed on 50km spatial windows (Eurostat density grid and OpenStreetMap)

}

\sframe{Network construction}{

%We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.

\justify

\textbf{Two layers: } population density (threshold $\theta_P$) and network characteristics (threshold $\theta_N$) taken among \{Number of edges, Number of vertices, Cyclomatic number $\mu$, Euclidian efficiency $v$ \}; percolated with a radius $r_0$

% banos2012towards

\bigskip

\textbf{Rationale: } \textit{two locations will be in relation if they are close, have a high population density and given network characteristics.}

% note : euclperf should be minimized whereas other maximized : issue ?

\bigskip

\textbf{Implementation: } construction of a single layer spatial network given the condition on the two layers and distances; extraction of connected components.

\bigskip

\textbf{Experience plan: } grid sampling for $r_0,\theta_P,\theta_N$ and network variables; additional gravity potential parameters $\gamma,d_0$ (detailed after)

$\rightarrow$ 4800 parameter points

%We systematically explore the clusters obtained for 840 parameter configurations.

% + explain openmole -> later in potential developments

}

\sframe{Results: endogenous mega-regions}{

%Maps reveal that most configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}.

\justify

\vspace{-0.4cm}
\textit{Extraction of endogenous polycentric mega-city regions \cite{hall2006polycentric}}

\medskip

\centering

% note : forgot r_0 in figs
\includegraphics[width=\textwidth,height=0.9\textheight]{figures/totalPop4183694_00056402_ecount850_radius8000.png}

}

\sframe{Different endogenous morphologies}{

% TODO for the paper : check effect of adding network threshold : methodo info on multilayer percol in itself

\hspace{-0.5cm}
\includegraphics[width=0.52\textwidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}
\includegraphics[width=0.52\textwidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}

}

\sframe{Characterizing sustainibility}{

% We use this endogenous definition of regional urban systems the percolation algorithm produced to evaluate their sustainability, in terms of conflicting objectives of economic integration and greenhouse gases emissions. Applying a gravity model to each region, we estimate transportation flows within each and extrapolate emissions by coupling with the Edgar emission database \citep{janssens2017edgar} and economic activities with a scaling law of population.

% explain indicators : convex hull.

\justify

\textbf{Application: } sustainability indicators for the endogenous urban regions; proxys for two conflicting dimensions: GHG emissions and economic integration \cite{viguie2012trade}.

\bigskip

\textbf{Data: } EDGAR database for GHG emissions (v4.3.2)

\cite{janssens2017edgar}

\bigskip

\textbf{Estimation: } Abstract flows approximated with a gravity model

\[
\phi_{ij} = \left(\frac{v_i v_j}{(\sum_k v_k)^2}\right)^\gamma \cdot \exp\left(\frac{-d_{ij}}{d_0}\right)
\]

where $v_k$ are either effective local GHG emissions or population (economic activity scaling law of population \cite{bettencourt2007growth})

\medskip

$\rightarrow$ sum of flows within the geographical span of the cluster (convex hull) approximate potential emissions and economic activity

}

\sframe{Pareto fronts for sustainability}{

%  We show therein that different population, network and distance thresholds yield different performances in terms of sustainability, exhibiting a Pareto front. This suggests policies in terms of regional integration to increase the sustainability of mega-city regions.

\textit{Superposing Pareto front for observed population and emissions, on all clusters.}

\centering

\medskip

\includegraphics[height=0.8\textheight]{figures/full_effective_pareto.png}

% -> full fronts in reserve slide

}

\sframe{Pareto fronts for aggregated indicators}{

\textit{Aggregated sustainability indicators suggest some configurations are more Pareto efficient (high $\gamma$ regime, activities with high added value).}

\centering

\medskip

\includegraphics[height=0.8\textheight]{figures/aggreg_paretos_radiuspopthq.png}

}

\sframe{An optimal morphology ?}{

%# Cumulative Proportion  0.7296 0.9650 0.99761 1.00000
%                PC1       PC2         PC3         PC4
%moran   -0.3088585 0.9493848 -0.04444327  0.03605266
%avgdist  0.5417362 0.1415668 -0.82239570 -0.10072759
%entropy  0.5108424 0.2140447  0.45847647 -0.69499942
%slope    0.5917502 0.1811415  0.33390034  0.71100630

\textit{Optimal degree of monocentricity of the system for emissions.}

\centering

\medskip

\includegraphics[height=0.85\textheight]{figures/aggreg_morpho_pc1-emissions.png}

}

\sframe{Morphological trade-offs}{

\vspace{-0.3cm}

\textit{No ``optimal'' cities, different forms yield different compromises in terms of relative indicators.}

\centering

\medskip

\includegraphics[height=0.85\textheight]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale.png}

}

\sframe{Towards a systematic calibration}{

% Further work will consist in the use of calibration heuristics \citep{reuillon2013openmole} to find in a more robust way optimal parameter values.

% -> depending on results

Grid sampling to explore regions rapidly limited

$\rightarrow$ towards the use of genetic algorithms on grid, made smooth with the OpenMOLE software \url{https://next.openmole.org/}

\centering

\smallskip

\includegraphics[height=0.35\textheight]{figures/openmole.png}

\smallskip

\raggedright\justify

\footnotesize \textit{OpenMOLE: (i) embed any model as a black box; (ii) transparent access to main High Performance Computing environments; (iii) model exploration and calibration methods.}

\bigskip

\textbf{Apply to the summer school !} \texttt{https://exmodelo.org/}

}

\sframe{Discussion}{

% This work paves the way towards ?

\justify

\vspace{-1cm}

\textbf{Implications}

$\rightarrow$ Transfer to policy making recommandations: Pareto-optimal configuration can be used for the planning of regional transportation networks, subsidies policies, etc.
\smallskip

$\rightarrow$

\bigskip

\textbf{Developments}

%

% detail method to extrapolate parameters of gravity flows

% interesting result : compare these emission with the synthetic ? (then just a refinment ?) -> yes, better. show new Pareto fronts with this fixed parameters for only emissions, and varying for economic ?

$\rightarrow$ Extrapolation of transportation flows

}

\sframe{Conclusion}{

\justify

\vspace{-1cm}

% rq : try in all conclusion to come back to these points, close to Banos commadments ?

$\rightarrow$ %First step towards systematic benchmarks and multi-modeling. \textbf{Need for more systematic model exploration.} % model coupling / benchmarking

\smallskip

$\rightarrow$ %Model integration and multi-scalarity ? \textbf{Need for more integrated models.}

\smallskip

$\rightarrow$ %Multiple perspectives on urban systems ? \textbf{Need for more interdisciplinarity.}

\bigskip

\footnotesize

\textbf{Related works}

Raimbault, J. (2018). Calibration of a density-based model of urban morphogenesis. PloS one, 13(9), e0203516.

\smallskip

Raimbault, J. (2018). An Urban Morphogenesis Model Capturing Interactions between Networks and Territories. \textit{Forthcoming in Mathematics or Urban Morphogenesis.} arXiv:1805.05195.

\smallskip

Raimbault, J. (2018). Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution des r{\'e}seaux de transport et des territoires (Doctoral dissertation, Université Paris 7 Denis Diderot). \url{https://halshs.archives-ouvertes.fr/tel-01857741}

\smallskip

\textbf{Open repository} at \texttt{https://github.com/JusteRaimbault/UrbanMorphology}\\\smallskip
\textbf{Acknowledgments}: thanks to the \textit{EGI} for access to the infrastructure.

}

\sframe{Reserve slides}{

\centering

\Large

\textbf{Reserve Slides}

}

\sframe{Results: effective emissions}{

\textit{Effective emissions exhibit a supralinear scaling of population}

\medskip

\centering

\includegraphics[height=0.9\textheight]{figures/aggreg_effective.png}

}

\sframe{Results: all clusters Pareto fronts}{

\textit{Variation of Pareto front patterns when potential parameter $\gamma,d_0$ vary.}

\medskip

\centering

\includegraphics[width=0.9\textwidth]{figures/full_paretos.png}

}

\sframe{Results: an optimal morphology}{

\textit{More monocentric areas are more optimal in terms of relative emissions and efficiency ?}

\medskip

\centering

\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relefficiency.png}
\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relemissions.png}

% TODO : curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.

}

%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{/Users/juste/ComplexSystems/CityNetwork/Biblio/Bibtex/CityNetwork,biblio}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}